---
title: "DA3_second_assignment"
author: "Akos Almasi"
date: "2023-02-09"
output:
  html_document:
    df_print: paged
---

```{r,message=FALSE, warning=FALSE}
# CLEAR MEMORY
rm(list=ls())
# Libraries used
library(tidyverse)
library(modelsummary)
library(ranger)
library(caret)
library(pdp)
library(gbm)
library(rattle)
library(tables)
```

## Data import
```{r}
area <- "Vienna"
data <- read.csv('https://raw.githubusercontent.com/akos-almasi/Data-Analysis-3/main/Second%20Assignment/listings.csv')

```

## Data exploration
```{r}
# Check the data
glimpse(data)

# Descriptive statistics
skimr::skim(data)

unique(data$property_type)
datasummary(property_type ~ Percent(), data = data, output = 'markdown')

Hmisc::describe(data$price)
```

## Data cleaning process
```{r}
# Focus on apartments 
data <- data %>% filter(room_type == 'Entire home/apt')

# Hosting 2-6 accommodates
data <- data %>% filter(accommodates >= 2 & accommodates <= 6)

# Change price column to numeric, remove dollar sign 
data$price <- as.numeric(as.character(gsub('\\$', '', data$price)))
# Change NA values to 0 in price column
data$price[is.na(data$price)] <- 0
# Filter prices that are 0
data <- data %>% filter(!price == 0)

# different type of property types
unique(data$property_type)
# Check how frequent each property type
datasummary(property_type ~ N + Percent(), data = data, output = 'markdown')

# Keep only the following property types: 
data <- data %>%
  filter(property_type %in% c('Entire condo', 'Entire loft', 'Entire home', 'Entire rental unit', 'Entire serviced apartment'))

# Extreme value
P01 <- function(x){quantile(x, 0.01)}
P99 <- function(x){quantile(x, 0.99)}
datasummary( price~ P0 + P01 + P25 + P50 + P75 +  P99 + P100 + Mean + SD,  
             data = data , output = 'markdown')
# Drop the unrealistically high values in price 
data <- data %>% filter(!price >= 380.42)
# Drop the unrealistically low values in price 
data <- data %>% filter(!price <= 27.00)

### BATHROOMS TEXT COLUMN ###
# remove bath from each element
data$bathrooms_text <- gsub(" bath","",data$bathrooms_text)
# remove s from each element
data$bathrooms_text <- gsub("s","",data$bathrooms_text)
# change 'Half-bath' to '0.5'
data$bathrooms_text <- gsub("Half-bath","0.5",data$bathrooms_text)
# Convert it to numeric create a new column called bathrooms
data$bathrooms <- as.numeric(data$bathrooms_text)
# Replace missing values with the most frequent value
frequent_value_bathrooms <- names(sort(table(data$bathrooms), decreasing = TRUE))[1]
data$bathrooms[is.na(data$bathrooms)] <- frequent_value_bathrooms
# Filter out the ones we have only 1 observation on
data <- data %>% filter(!bathrooms == 3.5)
data <- data %>% filter(!bathrooms == 4.5)

### BINARY COLUMNS ###
# Change t to 1 change f to 0 in binary columns
for (binary in c("host_is_superhost","host_has_profile_pic","host_identity_verified","has_availability","instant_bookable")){
  data[[binary]][data[[binary]]=="f"] <- 0
  data[[binary]][data[[binary]]=="t"] <- 1
}

### NEIGHBOURHOOD_CLEANSED COLUMN ###
# Change it to factor type
data$neighbourhood_cleansed <- as.factor(data$neighbourhood_cleansed)

### PROPERTY_TYPE COLUMN ###
# Change it to factor type
data$property_type <- as.factor(data$property_type)

###create days since first review ###
data <- data %>%
  mutate(days_since_review = as.numeric(as.Date(calendar_last_scraped,format="%Y-%m-%d") -
                                     as.Date(first_review ,format="%Y-%m-%d")))

sum(is.na(data$days_since_review))
### BEDS COLUMN ###
# Replace missing values with the most frequent value
frequent_value_beds <- names(sort(table(data$beds), decreasing = TRUE))[1]
data$beds[is.na(data$beds)] <- frequent_value_beds
# 
data <- data %>% filter (!beds == 7)

```

## Data creation from amenities column

The amenities column contains a lot of important information, decided to further investigate this column. Created different binary columns for each aminities that can be found in the dataset and summarized the most important them for each airbnb apartment. The following informations were considered important: tv, wifi, dishwasher, air conditioning, pool, free parking, paid parking, gym, heating, carbomonoxid alarm, smokealarm, elevator, indoor fireplace, shampoo, firstaid, oven, refrigerator, balcony, soundsystem, ironing, lockbox, securitycameras.
```{r}
############################ AMENITIES ##################################
# Clean amenities column
data$amenities<-gsub("\\{","",data$amenities)
data$amenities<-gsub("\\}","",data$amenities)
data$amenities<-gsub('\\"',"",data$amenities)

data$amenities<-as.list(strsplit(data$amenities, ","))
x <- levels(factor(unlist(data$amenities)))
data<-cbind(data,as.data.frame(do.call(rbind, lapply(lapply(data$amenities, factor, x), table))))

### TV ###
# Every column that contains TV in their name
d_tv_cols <- grep("tv", tolower(names(data)), fixed = TRUE)
# Create d_TV dummy variable, combine any of the columns that has 'tv' in their name and add it to data 
data <- data %>% mutate(d_TV = ifelse(rowSums(.[d_tv_cols]) > 0, 1, 0))

### WIFI ###
# Every column that contains wifi in their name
d_wifi_cols <- grep("wifi", tolower(names(data)), fixed = TRUE)
# Create d_wifi dummy variable, combine any of the columns that has 'wifi' in their name and add it to data
data <- data %>% mutate(d_wifi = ifelse(rowSums(.[d_wifi_cols]) > 0, 1, 0))

### DISH WASHER ###
# Every column that contains dishwasher in their name
d_dishwasher_cols <- grep("dishwasher", tolower(names(data)), fixed = TRUE)
# Create d_dishw dummy variable, combine any of the columns that has 'dishwasher' in their name and add it to data
data <- data %>% mutate(d_dishwash = ifelse(rowSums(.[d_dishwasher_cols]) > 0, 1, 0))

### AIR CONDITIONING ###
d_airconditioning_cols <- grep('air condition', tolower(names(data)), fixed = TRUE)
# Create d_airconditioning dummy variable, combine any of the columns that has 'air condition' in their name and add it to data
data <- data %>% mutate(d_airconditioning = ifelse(rowSums(.[d_airconditioning_cols]) > 0, 1, 0))

### POOL ###
# Remove whirlpool from each column name
colnames(data) <- gsub("whirlpool", "", colnames(data), ignore.case = TRUE)
# columns that contain pool
d_pool_cols <- grep('pool', tolower(names(data)), fixed = TRUE)
# Create d_pool dummy variable, combine any of the columns that has 'air condition' in their name and add it to data
data <- data %>% mutate(d_pool = ifelse(rowSums(.[d_pool_cols]) > 0, 1, 0))

### FREE PARKING ###
# columns that contain free parking
d_freeparking_cols <- grep('free parking', tolower(names(data)), fixed = TRUE)
# Create d_freeparking dummy variable, combine any of the columns that has 'free parking' in their name and add it to data
data <- data %>% mutate(d_freeparking = ifelse(rowSums(.[d_freeparking_cols]) > 0, 1, 0))

### PAID PARKING ###
# columns that contain paid parking
d_paidparking_cols <- grep('paid parking', tolower(names(data)), fixed = TRUE)
# Create d_paidparking dummy variable, combine any of the columns that has 'paid parking' in their name and add it to data
data <- data %>% mutate(d_paidparking = ifelse(rowSums(.[d_paidparking_cols]) > 0, 1, 0))

### GYM ###
# columns that contain gym
d_gym_cols <- grep('gym', tolower(names(data)), fixed = TRUE)
# Create d_gym dummy variable, combine any of the columns that has 'gym' in their name and add it to data
data <- data %>% mutate(d_gym = ifelse(rowSums(.[d_gym_cols]) > 0, 1, 0))

### HEATING ###
d_heating_cols <- grep('heating', tolower(names(data)), fixed = TRUE)
# Create d_heating dummy variable, combine any of the columns that has 'heating' in their name and add it to data
data <- data %>% mutate(d_heating = ifelse(rowSums(.[d_heating_cols]) > 0, 1, 0))

### CARBONMONOXID ###
d_carbonmonoxid_cols <- grep('monoxid', tolower(names(data)), fixed = TRUE)
# Create d_monoxid dummy variable, combine any of the columns that has 'monoxid' in their name and add it to data
data <- data %>% mutate(d_monoxid = ifelse(rowSums(.[d_carbonmonoxid_cols]) > 0, 1, 0))

### SMOKEALARM ###
d_smokealarm_cols <- grep('smoke', tolower(names(data)), fixed = TRUE)
# Create d_smokealarm dummy variable, combine any of the columns that has 'smoke' in their name and add it to data
data <- data %>% mutate(d_smokealarm = ifelse(rowSums(.[d_smokealarm_cols]) > 0, 1, 0))

### ELEVATOR ###
d_elevator_cols <- grep('elevator', tolower(names(data)), fixed = TRUE)
# Create d_elevator dummy variable, combine any of the columns that has 'elevator' in their name and add it to data
data <- data %>% mutate(d_elevator = ifelse(rowSums(.[d_elevator_cols]) > 0, 1, 0))

### INDOOR FIREPLACE ###
d_indoorfireplace_cols <- grep('indoor fireplace', tolower(names(data)), fixed = TRUE)
# Create d_indoorfireplace dummy variable, combine any of the columns that has 'indoor fireplace' in their name and add it to data
data <- data %>% mutate(d_indoorfireplace = ifelse(rowSums(.[d_indoorfireplace_cols]) > 0, 1, 0))

### SHAMPOO ###
d_shampoo_cols <- grep('shampoo', tolower(names(data)), fixed = TRUE)
# Create d_shampoo dummy variable, combine any of the columns that has 'shampoo' in their name and add it to data
data <- data %>% mutate(d_shampoo = ifelse(rowSums(.[d_shampoo_cols]) > 0, 1, 0))

### FIRST AID ###
d_firstaid_cols <- grep('first aid', tolower(names(data)), fixed = TRUE)
# Create d_firstaid dummy variable, combine any of the columns that has 'first aid' in their name and add it to data
data <- data %>% mutate(d_firstaid = ifelse(rowSums(.[d_firstaid_cols]) > 0, 1, 0))

### OVEN ###
d_oven_cols <- grep('oven', tolower(names(data)), fixed = TRUE)
# Create d_oven dummy variable, combine any of the columns that has 'first aid' in their name and add it to data
data <- data %>% mutate(d_oven = ifelse(rowSums(.[d_oven_cols]) > 0, 1, 0))

### REFRIGERATOR ###
d_refrigerator_cols <- grep('refrigerator', tolower(names(data)), fixed = TRUE)
# Create d_refrigerator dummy variable, combine any of the columns that has 'refrigerator' in their name and add it to data
data <- data %>% mutate(d_refrigerator = ifelse(rowSums(.[d_refrigerator_cols]) > 0, 1, 0))

### BALCONY ###
d_balcony_cols <- grep('balcony', tolower(names(data)), fixed = TRUE)
# Create d_balcony dummy variable, combine any of the columns that has 'balcony' in their name and add it to data
data <- data %>% mutate(d_balcony = ifelse(rowSums(.[d_balcony_cols]) > 0, 1, 0))

### SOUND SYSTEM ###
d_soundsystem_cols <- grep('sound system', tolower(names(data)), fixed = TRUE)
# Create d_soundsystem dummy variable, combine any of the columns that has 'sound system' in their name and add it to data
data <- data %>% mutate(d_soundsystem = ifelse(rowSums(.[d_soundsystem_cols]) > 0, 1, 0))

### IRON ###
d_iron_cols <- grep('iron', tolower(names(data)), fixed = TRUE)
# Create d_iron dummy variable, combine any of the columns that has 'iron' in their name and add it to data
data <- data %>% mutate(d_iron = ifelse(rowSums(.[d_iron_cols]) > 0, 1, 0))

### LOCK BOX ###
d_lockbox_cols <- grep('lockbox', tolower(names(data)), fixed = TRUE)
# Create d_lockbox dummy variable, combine any of the columns that has 'lockbox' in their name and add it to data
data <- data %>% mutate(d_lockbox = ifelse(rowSums(.[d_lockbox_cols]) > 0, 1, 0))

#### SECURITY ####
d_securitycameras_cols <- grep('security', tolower(names(data)), fixed = TRUE)
data <- data %>% mutate(d_securitycameras = ifelse(rowSums(.[d_securitycameras_cols]) > 0, 1, 0))

## Remove the columns that I generated from the amenities column
data <- data[, !(names(data) %in% x)]
```
## Data creation from description column

I realized that many people write the size of the apartment in this column, I wanted to check if I could find this information for most of the apartments. Unfortunately, this was not the case, some write in square feet instead of square meters, and some rows are missing this information completely. Overall it might be worth investigating this further, in my opinion it would be an important information for price forecasting. 
```{r, warning=FALSE}
### DESCRIPTION COLUMN ###
# Find the square meter of apartments in the description column 
data$description1 <- gsub("^.*?([0-9]+).*?(sqm|m2|m²|m ²|square meter|squaremeter).*$", "\\1", 
                          data$description)

data$description1 <- as.numeric(data$description1)
sum(is.na(data$description1))
```
## Drop unnecessary columns
```{r}
data <- data %>% select(-c('name','source','description','scrape_id',"listing_url", "last_scraped",
                           "listing_url", "picture_url", "host_id", "host_url",
                           'host_name','host_location','host_thumbnail_url', 
                           'host_picture_url', 'host_neighbourhood', 'host_about',"   (IKEA) refrigerator",
                           "  induction stove","  oven","   refrigerator", "  stainless steel oven", 
                           "last_review","first_review", "calendar_last_scraped", 
                           "calendar_updated", "calendar_last_scraped",'host_since',
                           'neighborhood_overview','neighbourhood', 'neighbourhood_group_cleansed',
                           'host_verifications','bathrooms_text', " Philipps   refrigerator", 'license', 
                           "  electric stove","  stainless steel gas stove", 'amenities'))

```


## Create train and holdout samples
```{r}
# set seed
set.seed(20230201)

train_indices <- as.integer(createDataPartition(data$price, p = 0.7, list = FALSE))
data_train <- data[train_indices, ]
data_holdout <- data[-train_indices, ]

# Check the number of observations in each
dim(data_train)
dim(data_holdout)
```

## Define models
```{r}
# Basic Variables incl. neighbourhood
basic_vars <- c(
  "accommodates", "beds","minimum_nights","bathrooms","maximum_nights",
  "property_type","neighbourhood_cleansed")

# reviews
reviews <- c("number_of_reviews")

# dummy variables
amenities <-  grep("^d_.*", names(data), value = TRUE)
 
# interactions for the LASSO
X1  <- c("accommodates*property_type","d_airconditioning*property_type", "maximum_nights*property_type", 'minimum_nights*property_type')
# with boroughs
X2  <- c("property_type*neighbourhood_cleansed", "accommodates*neighbourhood_cleansed")

predictors_2 <- c(basic_vars, reviews, amenities)
predictors_E <- c(basic_vars, reviews, amenities, X1,X2)

# do 5-fold CV
train_control <- trainControl(method = "cv",
                              number = 5,
                              verboseIter = FALSE)

```

## Random Forest model
```{r}
### set tuning ###
tune_grid <- expand.grid(
  # the number of predictors randomly sampled as candidates at each split. The value of 8 means that 
  # 8 predictors will be randomly sampled as candidates at each split.
  .mtry = c(8,10,12),
  # the rule for splitting the nodes in the decision tree. The value of "variance" means that the split is based on
  # the variance of the target variable.
  .splitrule = "variance",
  # the minimum size of the terminal nodes, or the minimum number of observations required in a terminal node. 
  # The value of 50 means that a terminal node must have at least 50 observations.
  .min.node.size = c(15,25,35)
)

### Random forest model ###
set.seed(1234)
system.time({
  rf_model <- train(
    formula(paste0("price ~", paste0(predictors_2, collapse = " + "))),
    data = data_train,
    method = "ranger",
    trControl = train_control,
    tuneGrid = tune_grid,
    importance = "impurity"
  )
})

rf_model

# DIAGNOSTICS 

# Variable Importance Plots 

# first need a function to calculate grouped varimp
group.importance <- function(rf.obj, groups) {
  var.imp <- as.matrix(sapply(groups, function(g) {
    sum(importance(rf.obj)[g], na.rm = TRUE)
  }))
  colnames(var.imp) <- "MeanDecreaseGini"
  return(var.imp)
}


rf_model_var_imp <- ranger::importance(rf_model$finalModel)/1000
rf_model_var_imp_df <-
  data.frame(varname = names(rf_model_var_imp),imp = rf_model_var_imp) %>%
  mutate(varname = gsub("f_neighbourhood_cleansed", "Borough:", varname) ) %>%
  mutate(varname = gsub("f_room_type", "Room type:", varname) ) %>%
  arrange(desc(imp)) %>%
  mutate(imp_percentage = imp/sum(imp))

rf_model_var_imp_df

# full varimp plot, top 10 only
ggplot(
    rf_model_var_imp_df[1:10,], 
    aes(x=reorder(varname, imp), y=imp_percentage)) +
  geom_point(color='skyblue', size=3) +
  geom_segment(
    aes(x=varname,xend=varname,y=0,yend=imp_percentage), 
    color='skyblue', size=1) +
  ylab("Importance (Percent)") +
  xlab("Variable Name") +
  labs(title = 'Top 10 variables in RF model') + 
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_bw()

# Partial dependence plots
# 1) number of accomodates
pdp_n_acc_rf <- pdp::partial(rf_model, 
                          pred.var = "accommodates", 
                          pred.grid = distinct_(data_holdout, "accommodates"), 
                          train = data_train)

pdp_n_acc_rf %>%
  autoplot( ) +
  geom_point(color='skyblue', size=2) +
  geom_line(color='skyblue', size=1) +
  ylab("Predicted price") +
  xlab("Accommodates (persons)") +
  labs(title = 'Partial dependence plot for number of accomodates (RF)') + 
  theme_bw()

# 2) property-type
pdp_n_property_rf <- pdp::partial(rf_model, 
                          pred.var = "property_type", 
                          pred.grid = distinct_(data_holdout, "property_type"), 
                          train = data_train)

pdp_n_property_rf %>%
  autoplot( ) +
  geom_point(color='skyblue', size=3) +
  ylab("Predicted price") +
  xlab("Property type") +
  labs(title = 'Partial dependence plot for property type (RF)') + 
  theme_bw()

```

## Gradient Boosting Model
```{r, warning=FALSE}
# set tuning
gbm_grid <-  expand.grid(interaction.depth = c(1, 5, 10), 
                          n.trees = (4:10)*50, 
                          shrinkage = 0.1, 
                          n.minobsinnode = 20)
                         
# The model         
set.seed(1234)
system.time({
  gbm_model <- train(formula(paste0("price ~", paste0(predictors_2, collapse = " + "))),
                     data = data_train,
                     method = "gbm",
                     trControl = train_control,
                     verbose = FALSE,
                     tuneGrid = gbm_grid)
})

gbm_model
gbm_model$finalModel

# Partial dependence plots
# 1) number of accomodates
pdp_n_acc_gbm <- pdp::partial(gbm_model, 
                          pred.var = "accommodates", 
                          pred.grid = distinct_(data_holdout, "accommodates"), 
                          train = data_train)

pdp_n_acc_gbm %>%
  autoplot( ) +
  geom_point(color='skyblue', size=2) +
  geom_line(color='skyblue', size=1) +
  ylab("Predicted price") +
  xlab("Accommodates (persons)") +
  labs(title = 'Partial dependence plot for number of accomodates (GBM)') + 
  theme_bw()

# 2) property-type
pdp_n_property_gbm <- pdp::partial(gbm_model, 
                          pred.var = "property_type", 
                          pred.grid = distinct_(data_holdout, "property_type"), 
                          train = data_train)

pdp_n_property_gbm %>%
  autoplot( ) +
  geom_point(color='skyblue', size=3) +
  ylab("Predicted price") +
  xlab("Property type") +
  labs(title = 'Partial dependence plot for property type (GBM)') + 
  theme_bw()

```

## OLS model

```{r, warning=FALSE}
set.seed(1234)
system.time({
  ols_model <- train(
    formula(paste0("price ~", paste0(predictors_2, collapse = " + "))),
    data = data_train,
    method = "lm",
    trControl = train_control
  )
})

ols_model_coeffs <-  ols_model$finalModel$coefficients
ols_model_coeffs_df <- data.frame(
  "variable" = names(ols_model_coeffs),
  "ols_coefficient" = ols_model_coeffs
) %>%
  mutate(variable = gsub("`","",variable))

# Partial dependence plots
# 1) number of accomodates
pdp_n_acc_ols <- pdp::partial(ols_model, 
                          pred.var = "accommodates", 
                          pred.grid = distinct_(data_holdout, "accommodates"), 
                          train = data_train)

pdp_n_acc_ols %>%
  autoplot( ) +
  geom_point(color='skyblue', size=2) +
  geom_line(color='skyblue', size=1) +
  ylab("Predicted price") +
  xlab("Accommodates (persons)") +
  labs(title = 'Partial dependence plot for number of accomodates (OLS)') + 
  theme_bw()

# 2) property-type
pdp_n_property_ols <- pdp::partial(ols_model, 
                          pred.var = "property_type", 
                          pred.grid = distinct_(data_holdout, "property_type"), 
                          train = data_train)

pdp_n_property_ols %>%
  autoplot( ) +
  geom_point(color='skyblue', size=3) +
  ylab("Predicted price") +
  xlab("Property type") +
  labs(title = 'Partial dependence plot for property type (OLS)') + 
  theme_bw()

ols_model
```

## Lasso model

```{r, warning=FALSE}
set.seed(1234)
system.time({
  lasso_model <- train(
    formula(paste0("price ~", paste0(predictors_E, collapse = " + "))),
    data = data_train,
    method = "glmnet",
    preProcess = c("center", "scale"),
    tuneGrid =  expand.grid(
      "alpha" = 1, 
      "lambda" = seq(0.01, 0.25, by = 0.01)),
    trControl = train_control
  )
})

lasso_model

# Coefficients
lasso_coeffs <- coef(
  lasso_model$finalModel,
  lasso_model$bestTune$lambda) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "variable") %>%
  rename(lasso_coefficient = `s1`)

# Coefficients that are not 0
lasso_coeffs_non_null <- lasso_coeffs[!lasso_coeffs$lasso_coefficient == 0,]
# Regression coefficients including: OLS model, Lasso model
regression_coeffs <- merge(ols_model_coeffs_df, lasso_coeffs_non_null, by = "variable", all=TRUE)




# Partial dependence plots
# 1) number of accomodates
pdp_n_acc_lasso <- pdp::partial(lasso_model, 
                          pred.var = "accommodates", 
                          pred.grid = distinct_(data_holdout, "accommodates"), 
                          train = data_train)

pdp_n_acc_lasso %>%
  autoplot( ) +
  geom_point(color='skyblue', size=2) +
  geom_line(color='skyblue', size=1) +
  ylab("Predicted price") +
  xlab("Accommodates (persons)") +
  labs(title = 'Partial dependence plot for number of accomodates (Lasso)') + 
  theme_bw()

# 2) property-type
pdp_n_property_lasso <- pdp::partial(lasso_model, 
                          pred.var = "property_type", 
                          pred.grid = distinct_(data_holdout, "property_type"), 
                          train = data_train)

pdp_n_property_lasso %>%
  autoplot( ) +
  geom_point(color='skyblue', size=3) +
  ylab("Predicted price") +
  xlab("Property type") +
  labs(title = 'Partial dependence plot for property type (Lasso)') + 
  theme_bw()

```

## Summary of the models

```{r}
# The final models all together
final_models <-
  list("OLS" = ols_model,
       "LASSO (model w/ interactions)" = lasso_model,
       "Random forest model" = rf_model,
       "GBM"  = gbm_model)


results <- resamples(final_models) %>% summary()


# Cross validated RMSE
df_CV_RMSE <- imap(
  final_models, 
  ~{mean(results$values[[paste0(.y,"~RMSE")]])}
) %>% 
  unlist() %>% 
  as.data.frame() %>%
  rename("CV RMSE" = ".")


# RMSE on the holdout data
df_holdout_RMSE <- map(
  final_models, 
  ~{RMSE(
    predict(.x, newdata = data_holdout), 
    data_holdout[["price"]]
  )}
) %>% 
  unlist() %>% 
  as.data.frame() %>%
  rename("Holdout RMSE" = ".")
# Create a key column to merge the 2 tables
df_CV_RMSE$models <- rownames(df_CV_RMSE)
df_holdout_RMSE$models <- rownames(df_holdout_RMSE)
# Merge the two tables
merged_RMSE <- merge(df_CV_RMSE, df_holdout_RMSE, by = 'models')
merged_RMSE
```


## Compare actual vs predicted on the holdout data

```{r}
# Predict on the holdout data
data_holdout$gbm_price <- predict(gbm_model, data_holdout)
data_holdout$rf_price <- predict(rf_model, data_holdout)

# Set colors for RF and GBM model model
colors <- c('GBM'='#990033', 'RF'= 'navyblue')

# Visualize the result
ggplot( 
  data_holdout, aes(x = price)   ) +
  geom_point( 
    aes(y = gbm_price, color = 'GBM' ), 
    size = 2 ) +
  geom_point( 
    aes(y = rf_price, color = 'RF' ), 
    size = 2 ) +
  geom_abline( 
    intercept = 0, 
    slope = 1, 
    size = 1, 
    linetype = 'dashed'
  ) +
  scale_color_manual(breaks = c('GBM', 'RF'), values = colors) + 
  xlim(-1,max(data_holdout$price))+
  ylim(-1,max(data_holdout$price))+
  labs(
    x='Actual price (US$)',
    y='Predicted price (US$)', 
    color = 'Models',
    title = "Predicted vs actual price - GBM & RF")+
  theme_bw() +
  theme(legend.position = 'left')


```




### Extra graphs for report

```{r}
# Airbnb price distribution
ggplot(data = data, aes(x = price)) + 
  geom_histogram(
    aes(y = (..count..)/sum(..count..)),
    bins = 30, fill = 'skyblue4', col = 'white', size = 0.25, alpha = 0.8) + 
  theme_bw() + 
  labs(
    title = "The distribution of airbnb apartment prices", 
    x = "Price (US dollar)", y = "percent of total"
  ) +
  scale_y_continuous(
    expand = c(0.01,0.01),
    labels = scales::percent_format(accuracy = 0.1)
  ) +
  theme(plot.title = element_text(size = rel(1))) +
  scale_x_continuous(expand = c(0.01,0.01))


# What kind of prices to expect on average for each property type, how many observations we have on each of them
data %>%   
  ggplot() +
  geom_jitter(aes(x = as.factor(property_type), y = price), width = 0.25, color = "skyblue4", size = 1)+
  geom_crossbar(data = data %>% group_by(property_type) %>% dplyr::summarize( avg = mean(price)), aes(x = as.factor(property_type), y = avg, ymin = avg, ymax = avg), size=0.5,col="orangered1", width = 0.55) +
  theme_bw() +
  scale_y_continuous(limits = c(0,400) , 
                     breaks = seq(0, 400, by=50)) +
  labs(y ='Price in USD',x = 'Property type') +
  ggtitle(paste0('Information on the price based on the property type'))



# Variable importance plot for the best performing model RF
rf_model_var_imp <- ranger::importance(rf_model$finalModel)/1000
rf_model_var_imp_df <-
  data.frame(varname = names(rf_model_var_imp),imp = rf_model_var_imp) %>%
  mutate(varname = gsub("f_neighbourhood_cleansed", "Borough:", varname) ) %>%
  mutate(varname = gsub("f_room_type", "Room type:", varname) ) %>%
  arrange(desc(imp)) %>%
  mutate(imp_percentage = imp/sum(imp))

rf_model_var_imp_df

# full varimp plot, top 10 only
ggplot(
    rf_model_var_imp_df[1:10,], 
    aes(x=reorder(varname, imp), y=imp_percentage)) +
  geom_point(color='skyblue4', size=3) +
  geom_segment(
    aes(x=varname,xend=varname,y=0,yend=imp_percentage), 
    color='skyblue4', size=1) +
  ylab("Importance (Percent)") +
  xlab("Variable Name") +
  labs(title = 'Variable importance plot for RF model') + 
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0,0.15)) +
  scale_x_discrete(labels = c('1 bathroom', '3 beds', '2 bathrooms', 'minimum nights', 'maximum nights', 'TV', 'air conditioning', 'number of reviews', 'Innere Stadt district', 'number of accomodates')) +
  theme_bw()


```





