---
title: "Final project web scraping"
author: "Akos"
date: "2022-12-27"
output:
  prettydoc::html_pretty:
    theme: architect
  html_document:
    df_print: paged
---

## Web scraping
#### Libraries used for the project
```{r, warning=FALSE, message=FALSE}
library(rvest)
library(httr)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(stringr)
```

#### Function to receive data from one house
```{r, eval=FALSE, warning=FALSE}
# URL of one house
url <- "https://realestatehungary.hu/ii-ker/elado+lakas/tegla-epitesu-lakas/33085768"
# Create an empty list
t_list <- list()
# Read html of a house
t <- read_html(url)
# Function
get_one_house <- function(url) {
  #Sys.sleep(30)
  t_list <- list()
  t <- read_html(url)
  t_list[["price"]]<- t %>% html_nodes('.text-nowrap span') %>% html_text()
  t_list[["area_sqm"]] <- t %>% html_nodes('.border-end .fs-5') %>% html_text()
  t_list[["room_count"]] <- t %>% html_nodes('.border-end+ .font-family-secondary .fs-5') %>% html_text()
  # Get the keys
  keys <- t %>% html_nodes('.ps-0:nth-child(1)') %>% html_text()
  keys <- trimws(keys)
  # Remove the first 5 since they are error messages
  keys <-  keys[-c(1,2, 49,50)]
  # Get the values
  values <- t %>% html_nodes(".fw-bold.ps-0") %>% html_text()
  values <- trimws(values)
  
  if (length(keys)==length(values)) {
    for (i in 1:length(keys)) {
      t_list[[keys[i]]] <- values[i]
    }
  }
  return(t_list)
}

# Check the function
t <- get_one_house("https://realestatehungary.hu/ii-ker/elado+lakas/tegla-epitesu-lakas/33085768")
```

#### get the link for one page
```{r, eval=FALSE, warning=FALSE}
get_one_page_link  <- function(url) {
  t <- read_html(url)
  all_link <- paste0("https://realestatehungary.hu", t %>% html_nodes('.listing__link') %>% html_attr('href'))
  house_link <- all_link[startsWith(all_link, 'https://realestatehungary.hu/ii-ker/elado+lakas/tegla-epitesu-lakas/')]
  return(house_link)
}
# Try the function
get_one_page_link("https://realestatehungary.hu/lista/elado+lakas+ii-ker?page=4")
```

```{r, eval=FALSE, warning=FALSE}
# The links for 40 pages
pages_for_link  <- paste0('https://realestatehungary.hu/lista/elado+lakas+ii-ker?page=', 1:40)

# All the house links for the first 40 pages
house_links <- sapply(pages_for_link,get_one_page_link)

# unlist
all_links <- as.character(unlist(house_links))
```

#### Save the data into lists, combine them and create a dataframe
```{r, eval=FALSE, warning=FALSE, message=FALSE}
# Save the data into lists
data_list1 <- lapply(all_links[1:50], get_one_house)
data_list2 <- lapply(all_links[51:100], get_one_house)
data_list3 <- lapply(all_links[101:150], get_one_house)
data_list4 <- lapply(all_links[151:200], get_one_house)
data_list5 <- lapply(all_links[201:250], get_one_house)
data_list6 <- lapply(all_links[251:300], get_one_house)
data_list7 <- lapply(all_links[301:350], get_one_house)
data_list8 <- lapply(all_links[351:400], get_one_house)
data_list9 <- lapply(all_links[401:450], get_one_house)
data_list10 <- lapply(all_links[451:500], get_one_house)
data_list11 <- lapply(all_links[501:550], get_one_house)
data_list12 <- lapply(all_links[551:600], get_one_house)
data_list13 <- lapply(all_links[601:650], get_one_house)
data_list14 <- lapply(all_links[651:700], get_one_house)
data_list15 <- lapply(all_links[701:750], get_one_house)
data_list16 <- lapply(all_links[751:771], get_one_house)
# combine together the lists
final_list <- c(data_list1,data_list2,data_list3, data_list4, data_list5,
                data_list6, data_list7, data_list8, data_list9, data_list10, data_list11, data_list12, data_list13,
                data_list14,data_list15,data_list16)
# create a data frame from our list
df <- rbindlist(final_list, fill = T)
# Save the data
saveRDS(df, file = "data.rds")
```
#### Or we can save the htmls and work with them instead
```{r, eval=FALSE, warning=FALSE, message=FALSE}
# save HTML files
i <- "https://realestatehungary.hu/ii-ker/elado+lakas/tegla-epitesu-lakas/33258083"
tail(strsplit(i, "/")[[1]], 1)

# Save html files into data folder
for (i in all_links) {
  
  download.file(i, destfile = paste0("/Users/akosalmasi/Desktop/CEU/Subjects/Web_scraping/Data/", 
                                     tail(strsplit(i, "/")[[1]], 1),".html"))
}

asd <- list.files("/Users/akosalmasi/Desktop/CEU/Subjects/Web_scraping/Data/", full.names = T)
```


#### Read the data
```{r,warning=FALSE, message=FALSE}
x2 <- readRDS(file = "data.rds")
```

## Data cleaning

#### area_sqm column
```{r,warning=FALSE, message=FALSE}
# Change it to numeric and remove m2 from the column
x2$area_sqm <- as.numeric(gsub("m2", "", x2$area_sqm))
```

#### price column
```{r,warning=FALSE, message=FALSE}
# Include only the ones that are in HUF
x2 <- x2 %>% filter(str_detect(price, ("million HUF")))
# remove million HUF from the column and rename it
x2$price_million_HUF <- gsub("million HUF", "", x2$price)
# change , to .
x2$price_million_HUF <- gsub(",", ".", x2$price_million_HUF)
# Change the column to numeric
x2$price_million_HUF <- as.numeric(x2$price_million_HUF)
# remove price column
x2 <- x2[,-1]
```

#### Balcony size column 
```{r,warning=FALSE, message=FALSE}
# Change it to numeric and remove m2
x2$balcony_m2 <- as.numeric(gsub("m2", "", x2$`Balcony size`))
# Remove the original column
x2 <- x2[,-15]
```

#### Drop empty columns
```{r,warning=FALSE, message=FALSE}
x2 <- x2 %>% select(-c(`error
            
    
        Connection error - we could not send the message. Please try again later.`, `check
            
    
        Your email was sent to the advertiser.`))
```

#### Condition of the real estate column 
```{r,warning=FALSE, message=FALSE}
# change not provided to NA
x2$`Condition of the real estate`[x2$`Condition of the real estate` == "not provided"] <- as.numeric(NA)
```

#### Building levels column
```{r,warning=FALSE, message=FALSE, echo=TRUE, results='hide'}
# Change not provided Building levels to NA
x2$`Building levels`[x2$`Building levels` == "not provided"] <- as.numeric(NA)
# Change the single-floor to 1
x2$`Building levels`[x2$`Building levels` == "single-floor"] <- 1
# Change it to integer
as.integer(x2$`Building levels`)
```

#### Orientation column
```{r}
x2$`Orientation`[x2$`Orientation` == "not provided"] <- as.numeric(NA)
```

#### Create a new price column that is in HUF
```{r}
x2$price_HUF <- x2$price_million_HUF * 1000000
```

#### Create a new column for price/square meter
```{r}
x2$price_per_sqm <- x2$price_HUF / x2$area_sqm
```

#### Create a new column for price categories
```{r}
x2 <- x2 %>% 
  mutate(
    # Price categories
    price_category = case_when(
      price_HUF <= 70000000 ~ "cheap", 
      price_HUF >70000000 & price_HUF <= 120000000 ~ "average",
      price_HUF > 120000000 & price_HUF <= 250000000 ~ "expensive",
      price_HUF > 250000000 ~ "luxurious"
    ))
```


## Visualizations

#### The price difference between the most expensive and the cheapest real estate
```{r,message=FALSE, warning=FALSE, fig.width=10, fig.height = 6, fig.align="center"}
# Most expensive real estates
expensive <- x2 %>% distinct(price_million_HUF) %>% top_n(3)
# Least expensive real estates
cheap <- x2 %>% distinct(price_million_HUF) %>% top_n(-3)
# Combine the two dataframes
together <- rbind(cheap, expensive)
# Create an ID column
together <- tibble::rowid_to_column(together, "ID")

ggplot(together , aes(x = reorder(ID, -price_million_HUF) , y = price_million_HUF)) + 
  geom_bar(stat="identity" ,  fill = 'skyblue4') +
  geom_segment(data = x2, aes(x = 0.5, xend = 6.5, y = mean(x2$price_million_HUF)
                               , yend = mean(x2$price_million_HUF)) , color = "orangered1") + 
  geom_label(aes(x = 1, y = mean(x2$price_million_HUF) + 25), 
             label = "Average price" ,
             size = 3.75, color = 'orangered1', fill = "white") +
  labs(y = 'Price in million HUF',x = 'Real estates') +
  scale_y_continuous(limits = c(0,800) , 
                     breaks = seq(0, 800, by=100)) +
  ggtitle(paste0('The price difference between the 3 most expensive and the 3 cheapest real estates')) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5 , size=14)) +
    theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```

#### Average price/ square meter
```{r,warning=FALSE, fig.width=10, fig.height = 6, fig.align="center"}
x2 %>%   
  ggplot() +
  geom_jitter(aes(x = as.factor(price_category), y = price_per_sqm), width = 0.25, color = "skyblue4", size = 1)+
  geom_crossbar(data = x2 %>% group_by(price_category) %>% dplyr::summarize( avg = mean(price_per_sqm)), aes(x = as.factor(price_category), y = avg, ymin = avg, ymax = avg), size=0.5,col="orangered1", width = 0.55) +
  theme_bw() +
    scale_y_continuous(limits = c(0,8500000) , 
                     breaks = seq(0, 8500000, by=500000)) +
  labs(y ='Price/square meter (HUF)',x = 'Price category') +
  ggtitle(paste0('Information on the price/ square meter based on the price category of the real estate')) +
  theme(plot.title = element_text(hjust = 0.5 , size=14)) +
  theme(axis.text.x = element_text(vjust= 1, hjust= 0.8, angle = 20)) +
  scale_x_discrete(limits = c("cheap", "average", "expensive", "luxurious"))

```

#### Correlation between price and area 
```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height = 6, fig.align="center"}
ggplot(x2, aes(x = area_sqm, y = price_million_HUF)) +
  geom_smooth(method = 'lm') +
  labs(y = 'Price (million HUF)',x = 'Area (square meter)') +
  geom_point() +
  theme_bw()
```

#### How much money you can ask on average based on the condition of your real estate
```{r,warning=FALSE, fig.width=10, fig.height = 6, fig.align="center"}
x2 %>%   
  ggplot() +
  geom_jitter(aes(x = as.factor(`Condition of the real estate`), y = price_million_HUF), width = 0.25, color = "skyblue4", size = 1)+
  geom_crossbar(data = x2 %>% group_by(`Condition of the real estate`) %>% dplyr::summarize( avg = mean(price_million_HUF)), aes(x = as.factor(`Condition of the real estate`), y = avg, ymin = avg, ymax = avg), size=0.5,col="orangered1", width = 0.55) +
  theme_bw() +
  scale_y_continuous(limits = c(0,800) , 
                     breaks = seq(0, 880, by=100)) +
  labs(y ='Price in million HUF',x = 'Condition of real estate') +
  ggtitle(paste0('Information on the price based on the condition of the real estate')) +
  theme(plot.title = element_text(hjust = 0.5 , size=14)) +
  theme(axis.text.x = element_text(vjust= 1, hjust= 0.8, angle = 20))
```

#### How much money you can ask on average based on the orientation of your real estate
```{r,warning=FALSE, fig.width=10, fig.height = 6, fig.align="center"}

x2 %>%   
  ggplot() +
  geom_jitter(aes(x = as.factor(`Orientation`), y = price_million_HUF), width = 0.25, color = "skyblue4", size = 1)+
  geom_crossbar(data = x2 %>% group_by(`Orientation`) %>% dplyr::summarize( avg = mean(price_million_HUF)), aes(x = as.factor(`Orientation`), y = avg, ymin = avg, ymax = avg), size=0.5,col="orangered1", width = 0.35) +
  theme_bw() +
  scale_y_continuous(limits = c(0,800) , 
                     breaks = seq(0, 800, by=100)) +
  labs(y ='Price in million HUF',x = 'Orientation') +
  ggtitle(paste0('Information on the price based on the orientation of the real estate')) +
  theme(plot.title = element_text(hjust = 0.5 , size=14)) +
  theme(axis.text.x = element_text(vjust= 1, hjust= 0.8, angle = 20))

```

#### Real estate price difference on average if you have an elevator or not
```{r,message=FALSE, warning=FALSE, fig.width=10, fig.height = 6, fig.align="center"}
# Check the average price if we have an elevator or not
elevator_matters <- x2 %>% group_by(Lift) %>% summarise(average_price_million_HUF = mean(price_million_HUF))
# Keep the ones where we have information about the elevator
elevator_matters <- elevator_matters[-c(2,4),]

ggplot(data = elevator_matters, aes(x = reorder(Lift, -average_price_million_HUF), y = average_price_million_HUF))+
  geom_bar(stat = "identity", fill = 'skyblue4', width = 0.4) +
    labs(y = 'Average price in million HUF',x = 'Elevator') +
    scale_y_continuous(limits = c(0,160) , 
                     breaks = seq(0, 160, by=20)) +
  theme_bw()

```
